---
title: "Class 11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Structural Bioinformatics

## Section 1: Introduction to the RCSB Protein Data Bank (PDB)

### PDB statistics

Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein?

```{r}
stats <- read.csv("PDB Data.csv")
stats
(stats$Total/sum(stats$Total))*100
sum(stats$Proteins)/sum(stats$Total)
```

## Section 3: Introduction to Bio3D in R

### Getting started with Bio3D

First we have to load the package we want to use

```{r}
library(bio3d)
```

Now we can use the functions from this bio3d package

### Working with individual PDB files

```{r}
pdb <- read.pdb("1hsg")
pdb
```

### Atom selection

```{r}
# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb,"calpha")
ca.inds
# Print details of the first few selected atoms
head(pdb$atom[ca.inds$atom,])
# Select chain A
a.inds <- atom.select(pdb,chain="A")
# Select C-alphas of chain A
ca.inds <- atom.select(pdb,"calpha",chain="A")
# We can combine multiple selection criteria to return their intersection
cab.inds <- atom.select(pdb,elety=c("CA","CB"),chain="A",resno=10:20)
```

We want to select out the protein and drug only parts of these molecular PDB files. 

```{r}
prot.inds <- atom.select(pdb,"protein")
pdb$atom[prot.inds$atom,]
prot.pdb <- trim.pdb(pdb,prot.inds)
write.pdb(prot.pdb,file="protein.pdb")
```

Lets do the same thing for our drug

```{r}
ligs.inds <- atom.select(pdb, "ligand")
lig.pdb <- trim.pdb(pdb,ligs.inds)
write.pdb(lig.pdb,file="ligand.pdb")
```

```{r}
install.packages("devtools")
devtools::install_bitbucket("Grantlab/bio3d-view")
library("bio3d.view")
pdb <- read.pdb("5p21")
view(pdb, "overview", col="sse")
```
