---
title: "Class 14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Asthma SNPs in MXL Population

Read file from Ensemble

http://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=362656008#373531_tablePanel

```{r}
mxl <- read.csv("MXL.csv")
head(mxl)
```

How many of each genotype are in this set?

```{r}
table(mxl$Genotype..forward.strand.)
```

Percent Values

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl)*100
```

### RNA-Seq

Fastq quality scores

Q: Does the first sequence have good quality? 

```{r}
q <- "DDDDCDEDCDDDDBBDDDCC@"
library(seqinr)
library(gtools)
asc(s2c(q))-33
```

### Population scale analysis

Does the genotype for rs9067378 influence expression of the important asthma linked genes we found?

```{r}
geno <- read.table("Expression.txt")
head(geno)
table(geno$geno)
```

```{r}
inds.gg <- geno$geno == "G/G"
summary(geno$exp[inds.gg])
inds.aa <- geno$geno == "A/A"
summary(geno$exp[inds.aa])
inds.ag <- geno$geno == "A/G"
summary(geno$exp[inds.ag])
```

Some nice figure that shows the distribution of these expression values?

```{r}
boxplot(exp ~ geno,geno)
```

