---
title: "Class 13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1: In silico docking of drugs to HIV-1 protease

### Obtaining and inspecting our input structure

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
hiv <- read.pdb(file.name)
hiv
```

Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure?
[HOH (127), MK1 (1)]
198 protein atoms residues / C alpha atoms
0 nucleic acid atoms residues / phosphate atoms
128 non-protein / nucleic atoms residues

### Prepare initial protein and ligand input files

```{r}
prot <- trim.pdb(hiv,"protein")
lig <- trim.pdb(hiv,"ligand")
write.pdb(prot,file="1hsg_protein.pdb")
write.pdb(lig,file="1hsg_ligand.pdb")
```

### Using AutoDockTools to setup protein docking input

Q2: Can you locate the binding site visually? Note that
crystal structures normally lack hydrogen atoms, why? 
Yes, I can visually locate the binding site.
Hydrogens are too small to be seen in normal crystal structures. 

Q3: Look at the charges. Does it make sense (e.g. based on your knowledge of the
physiochemical properties of amino acids)?
Yes, the charges make sense.

### Prepare the ligand

### Prepare a docking configuration file

## Section 2: Docking ligands into HIV-1 protease