---
title: "Class 06 Homework"
author: "Rachel Goodridge"
date: "January 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lecture 6 Hands-on Lab Sheet

Section 1: Improving Analysis Code by Writing Functions

Part B: Install the bio3d package using the command **install.packages("bio3d")**
```{r}
library(bio3d)
```

### The Original Code

Any copy/paste errors have been corrected. 
```{r}
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

### My Code

Question 6

Create a function to simplify and avoid errors.
```{r}
PDB_plot <- function(x) {
  chain <- trim.pdb(read.pdb(x), chain="A", elety="CA")
  plotb3(chain$atom$b, sse=chain, typ="l", ylab="Bfactor")
}
```
The input (x) will be the protein (in this case: "4AKE", "1AKE", or "1E4Y"). The function will read the protein PDB data and use this to create a plot for that protein. To call the function, type **PDB_plot(x)**.

Test to ensure the function works correctly.
```{r}
PDB_plot("4AKE")
PDB_plot("1AKE")
PDB_plot("1E4Y")
```

The plots created using this function match the plots created by the original code. Thus, the function has done it's job!

