---
title: "Class 6"
author: "Rachel Goodridge"
date: "January 24, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Section 1: Reading files again :-)

Here we are going to try using the **read.table()** function and friends to read some example flat files.

First lets try adding a simple plot

```{r}
plot(1:10)
```

Back to file reading...

```{r}
read.table("https://bioboot.github.io/bimm143_W19/class-material/test1.txt",header=TRUE,sep=",")
```

For this common CSV format we can use **read.csv()**

```{r}
file1 <- "https://bioboot.github.io/bimm143_W19/class-material/test1.txt"
data1 <- read.csv(file1)
data1
```
```{r}
file2 <- "https://bioboot.github.io/bimm143_W19/class-material/test2.txt"
data2 <- read.table(file2,header=TRUE,sep="$")
data2
```
```{r}
file3 <- "https://bioboot.github.io/bimm143_W19/class-material/test3.txt"
data3 <- read.table(file3)
data3
```

### Section 2: R Functions


My first silly function

```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

Lets try using this function

```{r}
add(1)
```

```{r}
add(1,100)
```

```{r}
add( c(1, 2, 3) )
```

```{r}
add(c(1, 2, 3),4)
```

```{r}
#add(1, 2, 2)
```

```{r}
#add(x=1, y=“b”)
```

When you find yourself doing the same thing three or more times, make a function instead.

Instead of this:
```{r}
#df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
#df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
#df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
#df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))
```

Make a function:
```{r}
rescale <- function(x) {
  rng <- range(x)
  (x-rng[1])/(rng[2]-rng[1])
}
```

Test on a small example where we know what the answer should be!
```{r}
rescale(1:10)
```
```{r}
#How would you get this function to run?
#rescale(c(1,2,NA,3,10))
```
```{r}
#What should your function do here?
#rescale(c(1,10,"string"))
```

Lets make a new version of the function to get rid of the NA error.
```{r}
rescale2 <- function(x,na.rm=TRUE) {
  rng <- range(x,na.rm=na.rm)
  (x-rng[1])/(rng[2]-rng[1])
}
```
```{r}
rescale2(c(1,2,NA,3,10))
```

Version 3 of the same rescale function with plot
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```
```{r}
rescale3(c(1:6,NA,10),plot=TRUE)
```

### Working with the bio3d package

To install this package I used the command **install.packages("bio3d")**

Lets get using it.

```{r}
library(bio3d)
```

```{r}
# Read a PDB file from the database
s1 <- read.pdb("4AKE") # kinase with drug
s1
```

Try out the provided code...
```{r}
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

